### 2018-03-12 draft.tex
WIP draft.pdf + Keeping up-to-date for PC work

\documentclass{scrartcl}

\usepackage[hidelinks]{hyperref}
\usepackage[none]{hyphenat}
\usepackage{setspace}
%\doublespace Plz no, I need to proofread this

\title{Which C++ architecture(s) could best support QoE-oriented map streaming in a player-building-based multiplayer game?}
\subtitle{COMP130 - Computing Architecture}
\date{\today}
\author{1707981}

\begin{document}
\maketitle
\pagenumbering{arabic}

\begin{abstract}
An exploration into QoE techniques and their compatibility with various softwrae architecture reveals something that isn't revealed until I finish this flipping essay. Also, QoS is linked to QoE is linked to QoP is linked to Soviet Russia.
\end{abstract}

\section{Proposal}

\section{Challenges of low bandwidth}

\section{Notes}
\subsection{China market findings}
China is in fact the most profitable games market as of 2017 \cite{https://technology.ihs.com/593405/china-largest-games-market-continues-its-growth-story}, yet the Internet speeds lag behind (figuratively speaking) at 7.6Mbps. Another country notorious for its online gaming community is Brazil, notorious for its unreliable Internet speeds, which averaged about the same--6.8Mbps--in 2017.\cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/q1-2017-state-of-the-internet-connectivity-report.pdf}. While a marked improvement over time is evident (doubling since 2015 \cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/akamai-state-of-the-internet-report-q1-2015.pdf}), to the extent that low-bandwidth games such as FPSes are comfortably playable, and the player reputation as in \cite{} is reduced.

This presents a heavy roadblock for MMO developers

Internet speeds in Brazil are notoriously low, which is remarkable for the fifth top source of DDoS attacks in Q1 201\cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/q4-2017-state-of-the-internet-security-report.pdf}

\subsection{Link between computer specs and Internet speeds}

\nocite{*}
\bibliography{references} 
\bibliographystyle{ieeetr}

\end{document}